---
title: "Eclust Vignette"
author: "Sahir R. Bhatnagar"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_height: 8
    fig_width: 11
    keep_md: yes
    toc: yes
    toc_depth: 4
    toc_float: no
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Load Required Packages

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
library(knitr)
library(data.table)
library(magrittr)
library(ggplot2)
library(latex2exp)
library(dplyr)
library(plyr)
library(glmnet)
library(stringr)
library(DT)
```


# Simulate Data

```{r data}

set.seed(123456)

# number of predictors
p = 10 

# number of test subjects
n = 200

# correlation between X's
rho = 0.50

# signal to noise ratio
signal_to_noise_ratio = 4

# names of the main effects, this will be used in many of the functions
main_effect_names <- paste0("x",1:p) 

# names of the active set
true_var_names <- c("x1","x2","x3","x4","x1:x2", "x1:x3", "x1:x4", "x2:x3", "x2:x4", "x3:x4")

# different true coefficient vectors as in Table 1 of Choi et al. 
beta1 <- c(7,2,1,1,0,0,0,0,0,0) %>% magrittr::set_names(true_var_names)
beta2 <- c(7,2,1,1,1,0,0,0.5,0.4,0.1) %>% magrittr::set_names(true_var_names)
beta3 <- c(7,2,1,1,7,7,7,2,2,1) %>% magrittr::set_names(true_var_names)
beta4 <- c(7,2,1,1,14,14,14,4,4,2) %>% magrittr::set_names(true_var_names)
beta5 <- c(0,0,0,0,7,7,7,2,2,1) %>% magrittr::set_names(true_var_names)

# simulate Toeplitz like correlation structure between X's
H <- abs(outer(1:p, 1:p, "-"))
cor <- rho^H

# generate X's from multivariate normal and label the matrix
DT <- MASS::mvrnorm(n = n, mu = rep(0,p), Sigma = cor) %>% 
    magrittr::set_colnames(paste0("x",1:p)) %>% 
    set_rownames(paste0("Subject",1:n))

# create X matrix which contains all main effects and interactions
# but not the intercept
X <- model.matrix(
    as.formula(paste0("~(",paste0(main_effect_names, collapse = "+"),")^2-1")), 
    data = DT %>% as.data.frame()) 

# generate response with user defined signal to noise ratio 
y.star <- X[,names(beta4)] %*% beta4
error <- rnorm(n)
k <- sqrt(var(y.star)/(signal_to_noise_ratio*var(error))) 
Y <- y.star + k*error 
colnames(Y) <- "Y"

# names of interaction variables assuming interaction terms contain a ":"
# this will be used in many of the functions
# names must appear in the same order as X matrix
interaction_names <- grep(":", colnames(X), value = T)
main_effect_names <- setdiff(colnames(X), interaction_names)
```

```{r results='asis', echo=FALSE}
DT::datatable(X)
```


# Analysis

## Running the Strong heredity interaction model once using the `shim` function

NOTE: if `lambda.beta = NULL` and `lambda.gamma = NULL` then this function will use a grid of tuning parameters based on `nlambda.beta` only and the `nlambda.gamma` parameter is ignored. Therefore, if for example `nlambda.beta = 10` then you must specify `nlambda = 10x10 = 100` otherwise the function will break.

```{r, eval=TRUE}
# load eclust library
library(eclust)

res <- shim(x = X, y = Y,
            main.effect.names = main_effect_names,
            interaction.names = interaction_names,
            lambda.beta = NULL,
            lambda.gamma = NULL, 
            nlambda.beta = 10,
            nlambda = 100, 
            threshold = 1e-04, 
            max.iter = 200,
            initialization.type = "ridge", 
            center = TRUE,
            normalize = TRUE, 
            verbose = TRUE, 
            cores = 2)

names(res)
res

```

### Main effect ($\beta$) parameter estimates


```{r, results='asis', echo=FALSE}
DT::datatable(round(as.matrix(res$beta),2), options = list(pageLength = 100))
```


### Interaction effect ($\alpha$) parameter estimates


```{r, results='asis', echo=FALSE}
DT::datatable(round(as.matrix(res$alpha),2), options = list(pageLength = 100))
```


### Sequence of Tuning Parameters

```{r, results='asis', echo=FALSE}
DT::datatable(res$tuning.parameters, options = list(pageLength = 100))
```



## Cross Validation using the `cv.shim` function

```{r, eval=TRUE}

library(doMC)
registerDoMC(cores = 10)
cv.res <- cv.shim(x = X, y = Y,
            main.effect.names = main_effect_names,
            interaction.names = interaction_names,
            lambda.beta = NULL,
            lambda.gamma = NULL,
            nlambda.beta = 10,
            nlambda = 100, 
            parallel = TRUE,
            type.measure = c("mse"), 
            nfolds = 5)

names(cv.res)
```

### Cross Validation Plot

```{r}
plot(cv.res)
```


### Coefficient Estimates

```{r}
coef(cv.res, s = "lambda.1se")
coef(cv.res, s = "lambda.min")
```





